---
title: "Exploring the data biases"
author: 
- Petra Hermankova^[Aarhus University, Denmark, https://orcid.org/0000-0002-6349-0540]
date: "`r format(Sys.Date())`"
output:
  html_document:
    theme: united
    toc: yes
    toc_float: true
    number_sections: true
    toc_depth: 3
    df_print: paged
---

# Initial setup

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(sf)
library(sfarrow)
library(tidyverse)
library(jsonlite)
library(tidyr)
library(ggplot2)
library(viridis)
```

```{r loading data}
getwd()
#LISTjson <- fromJSON("../../data/large_data/LIST_occupsorgs_industry_simple_20231206.json") # without people
#LISTjson <- fromJSON("../../data/large_data/LIST_occupsorgs_industry_people_20231206.json") # with people, HISCO cats
LISTjson <- fromJSON("../../data/large_data/LIST_occupsorgs_industry_harris_people_20231206.json") # with people, Harris cats

LIST <- as_tibble(LISTjson)
```

## Making data sample
```{r}
LISTsmall <- LIST %>% 
  #select(LIST.ID, occups, occups_N, occups_cats, type_of_inscription_auto, urban_context, urban_context_city, urban_context_pop_est, Longitude, Latitude, not_after, not_before, clean_text_interpretive_word, status_notation, people) %>% 
  select(`LIST-ID`, occups, occups_cats, people)

LISTsmalljson <- jsonlite::toJSON(LISTsmall, auto_unbox=TRUE)

write(LISTsmalljson, file="LISTsmall.json")

```

## Getting the status and occupational data out nested dataset

### Occupations
```{r}
# Matteo's solution 
LISTsample <- read_json("LISTsmall.json")

# initialise df with LIST-ID and occupations
LISTsample_occups_df <- NULL
for (i in 1:length(LISTsample)) {
  # per each element in the list find the string that contains info on the occups
  LISTsample_occups_i <- LISTsample[[i]]$occups
  # in some cases the string is empty, skip these and consider only non-empty ones
  if (!is_empty(LISTsample_occups_i)){
    # transform into a df
    LISTsample_occups_i_df <- unlist(LISTsample_occups_i)
    
    # attach LIST-ID
    LISTsample_occups_i_df <- data.frame(cbind(LISTsample[[i]]$`LIST-ID`,
                                               LISTsample_occups_i_df)) %>% 
      setNames(c("LIST-ID","occups"))
    
    #add to the previous iteration of the dataframe of LIST-IDs and occups
    LISTsample_occups_df <- rbind(LISTsample_occups_df, LISTsample_occups_i_df) 
  }
  #print(i)
}
```

```{r}
#save the results
write.csv(LISTsample_occups_df, "LIST_occups_all.csv",row.names = FALSE)
```

### Occupational categories
```{r}
# initialise df with LIST-ID and occupations
LISTsample_occups_df_harris <- NULL
for (i in 1:length(LISTsample)) {
  # per each element in the list find the string that contains info on the occups cats
  LISTsample_occups_ih <- LISTsample[[i]]$occups_cats
  # in some cases the string is empty, skip these and consider only non-empty ones
  if (!is_empty(LISTsample_occups_ih)){
    # transform into a df
    LISTsample_occups_ih_df <- unlist(LISTsample_occups_ih)
    
    # attach LIST-ID
    LISTsample_occups_ih_df <- data.frame(cbind(LISTsample[[i]]$`LIST-ID`,
                                               LISTsample_occups_ih_df)) %>% 
      setNames(c("LIST-ID","occups_cats"))
    
    #add to the previous iteration of the dataframe of LIST-IDs and occups
    LISTsample_occups_df_harris <- rbind(LISTsample_occups_df_harris, LISTsample_occups_ih_df) 
  }
  #print(i)
}
```

```{r}

#save the results
write.csv(LISTsample_occups_df_harris, "LIST_harris_all.csv",row.names = FALSE)

```

```{r}

# merge the two df by LIST ID

all_occups<- left_join(LISTsample_people_df, LISTsample_occups_df_harris, by = "LIST-ID")
LIST_all_people_harris<- left_join(all_occups, LISTsample_occups_df, by = "LIST-ID")

```

### People and their status

```{r}
# initialise df with LIST IDs and people  
LISTsample_people_df <- NULL
for (i in 1:length(LISTsample)) {
  # per each element in the list find the string that contains info on the people
  LISTsample_people_i <- LISTsample[[i]]$people
  # in some cases the string is empty, skip these and consider only non-empty ones
  if (!LISTsample_people_i == "None" && !LISTsample_people_i == "[]"){
    # we need to handle some irregularities with the string before reading it
    
    # delete illegal character sequences that mess up the json
    LISTsample_people_i <- str_replace_all(LISTsample_people_i,"\\\\",'\\') # corrects issue with double \\
    LISTsample_people_i <- str_replace_all(LISTsample_people_i,"'\\.","\\.") # corrects issue with '. 
    
    # change from Python dictionary to JSON syntax
    LISTsample_people_i <- str_replace_all(LISTsample_people_i,"None","null")
    LISTsample_people_i <- str_replace_all(LISTsample_people_i,"\n",",")
    LISTsample_people_i <- str_replace_all(LISTsample_people_i,"'",'"')
    
    # transform into a df
    LISTsample_people_i_df <- fromJSON(LISTsample_people_i)
    
    # attach LIST-ID
    #LISTsample_people_i_df <- cbind(`LIST-ID` = LISTsample[[i]]$`LIST-ID`, 
    #                                LISTsample_people_i_df)
    LISTsample_people_i_df <- cbind(`LIST-ID` = as.character(LISTsample[[i]]$`LIST-ID`),
                                    LISTsample_people_i_df)
    
    #add to the previous iteration of the dataframe of LIST-IDs and people
    LISTsample_people_df <- rbind(LISTsample_people_df,LISTsample_people_i_df) 
  }
  #print(i)
}


```
### Merging data together

```{r}

#save the results
write.csv(LISTsample_people_df,"LISTsample_people.csv",row.names = FALSE)

# merge together this info
# warning: this df is not in an ideal format: it has info in different rows 
LISTsample_df <- merge(LISTsample_occups_df, LISTsample_people_df)

write.csv(LISTsample_df,"LISTsample.csv",row.names = FALSE)

```

```{r}
# inscriptions with status
status <- LISTsample_people_df %>% 
  filter(!is.na(status))

# inscriptions with occupations
occups <- LISTsample_occups_df

all<- left_join(occups, LISTsample_people_df, by = "LIST-ID")

status <- all %>% 
  filter(!is.na(status))


status_occups<- status %>% 
  count(status, occups)

occups_status<- status %>% 
  count(occups, status)

write.csv(status_occups,"LIST_status_occups_overview.csv", row.names = TRUE)
write.csv(occups_status,"LIST_occups_status_overview.csv", row.names = TRUE)
```



## Exploration

### Occupational categories by status
```{r}
# explore  Harris categories by people's status

LIST_all_people_harris %>% 
  select(status, occups_cats) %>% 
  filter(!is.na(occups_cats)) %>% 
  count(occups_cats, status, sort = TRUE) -> harris_by_status
```

```{r}
harris_by_status
```

```{r}
# ONLY people with status and occupational categories

LIST_all_people_harris %>% 
    filter(!is.na(occups_cats)) %>% 
  filter(!is.na(status)) -> LIST_occups_and_status

  
```

```{r}
LIST_all_people_harris %>% 
  select(status, occups_cats) %>% 
  filter(!is.na(occups_cats)) %>% 
  filter(!is.na(status))%>% 
  mutate(status_clean = str_replace_all(string = status, pattern = "\\?", replacement = "")) %>% 
  ggplot(aes(y=status_clean, fill = occups_cats)) +
           geom_bar(position =  "dodge") +
theme_linedraw(base_size = 6) +
  labs(x = "Ratio", y = "Status on inscriptions", title = "Status of occupational groups (based on the EDH dataset)", subtitle = ggtitle(paste("n =", nrow(LIST_occups_and_status), "inscriptions with occupation and status"))) 

ggsave(width = 10, height = 6, dpi = 300, filename = "Harris_groups_by_status.png")
  
```

```{r}
LIST_all_people_harris %>% 
  select(status, occups_cats) %>% 
  filter(!is.na(occups_cats)) %>% 
  filter(!is.na(status))%>% 
  mutate(status_clean = str_replace_all(string = status, pattern = "\\?", replacement = "")) %>% 
  ggplot(aes(y=status_clean, fill = occups_cats)) +
           geom_bar(position =  "fill") +
theme_linedraw(base_size = 6) +
  labs(x = "Ratio", y = "Status on inscriptions", title = "Status of occupational groups (based on the EDH dataset)", subtitle = ggtitle(paste("n =", nrow(LIST_occups_and_status), "inscriptions with occupation and status"))) 

ggsave(width = 10, height = 6, dpi = 300, filename = "Harris_groups_by_status_ratio.png")
  
```

Commentary:
slaves do a little bit of everything, but are well underrepresented
freedmen services and managerial, food production


#### Military personnel
```{r}
military_personnel <- LIST_occups_and_status %>% 
  filter(status == str_subset(status, "military"))

# exploring military personnel, #ratio of all people that have both status an occupation mentioned in the inscription and ratio of people within the occupational category
LIST_all_people_harris %>% 
  filter(status=="military personnel") %>% 
  filter(!is.na(occups_cats)) %>% 
  count(occups_cats, sort = TRUE) %>% 
mutate(ratio_all = format(n/(nrow(LIST_occups_and_status)/100), scientific=FALSE) ) %>% mutate(ratio_category = format(n/(nrow(military_personnel)/100), scientific=FALSE) )

```

```{r}
LIST_all_people_harris %>% 
  filter(status=="military personnel") %>% 
  filter(!is.na(occups_cats)) %>% 
  count(occups, sort = TRUE)

### 1690 instances of cornicen = a hornblower, trumpeter, cornet-player
### 722 instances of clavicularius = A bar, bolt of the door maker, or guardian of a prison


```

#### Freedmen

```{r}
freedmen <- LIST_occups_and_status %>% 
  filter(status == str_subset(status, "freedmen"))


# exploring freedmen
LIST_all_people_harris %>% 
  filter(status=="freedmen / freedwomen") %>% 
  filter(!is.na(occups_cats)) %>% 
  count(occups_cats, sort = TRUE) %>% 
  mutate(ratio_all = format(n/(nrow(LIST_occups_and_status)/100), scientific=FALSE) ) %>% mutate(ratio_category = format(n/(nrow(freedmen)/100), scientific=FALSE) )
```

### Augustales
```{r}

augustales <- LIST_occups_and_status %>% 
  filter(status == str_subset(status, "Augustales"))

# exploring Augustales
LIST_all_people_harris %>% 
  filter(status=="Augustales") %>% 
  filter(!is.na(occups_cats)) %>% 
  count(occups_cats, sort = TRUE) %>% 
  mutate(ratio_all = format(n/(nrow(LIST_occups_and_status)/100), scientific=FALSE) ) %>% mutate(ratio_category = format(n/(nrow(augustales)/100), scientific=FALSE) )
```

#### Slaves

```{r}

slaves <- LIST_occups_and_status %>% 
  filter(status == str_subset(status, "slaves"))


# exploring slaves
LIST_all_people_harris %>% 
  filter(status=="slaves") %>% 
  filter(!is.na(occups_cats)) %>% 
  count(occups_cats, sort = TRUE) %>% 
  mutate(ratio_all = format(n/(nrow(LIST_occups_and_status)/100), scientific=FALSE) ) %>% mutate(ratio_category = format(n/(nrow(slaves)/100), scientific=FALSE) )
```

### Senators

```{r}

senators <- LIST_occups_and_status %>% 
  filter(status == str_subset(status, "senatorial"))

# exploring senators
LIST_all_people_harris %>% 
  filter(status=="senatorial order") %>% 
  filter(!is.na(occups_cats)) %>% 
  count(occups_cats, sort = TRUE) %>% 
  mutate(ratio_all = format(n/(nrow(LIST_occups_and_status)/100), scientific=FALSE) ) %>% mutate(ratio_category = format(n/(nrow(senators)/100), scientific=FALSE) )
```

### Elites and non-elites

What proportion of work was claimed by elites and nonelites.

```{r}

elites <- c("equestrian order", "decurial order, higher local offices", "senatorial order", "equestrian order", "military personnel", "lower local offices, administration of imperial estates", "emperor / imperial household", "rulers (foreign)")
non_elites <- c("freedmen / freedwomen", "slaves", "Augustales" )

elites_data <- LIST_occups_and_status %>% 
  filter(status %in% elites) %>% 
  count(occups_cats)

nonelites_data <- LIST_occups_and_status %>% 
  filter(status %in% non_elites) %>% 
  count(occups_cats)

nonelites_data[is.na(nonelites_data)] <- 0
```

```{r}

elites_nonelites<- left_join(elites_data, nonelites_data, by = "occups_cats") 

rename(elites_nonelites, elites_n = n.x) -> elites_nonelites
rename(elites_nonelites, nonelites_n = n.y) -> elites_nonelites

elites_nonelites<- elites_nonelites %>% 
  mutate(elites_ratio_per_cat = elites_n/(sum(elites_n)/100)) %>% 
  mutate(nonelites_n = if_else(is.na(nonelites_n), 0, nonelites_n)) %>% 
  mutate(nonelites_ratio_per_cat = nonelites_n/(sum(nonelites_n)/100))

```

```{r}
#save the results
write.csv(elites_nonelites, "LIST_occups_cats_by_status_elites_nonelites.csv",row.names = TRUE)
```

```{r}
# People mentioned on inscriptions but without extra filtering of the occupation (so there are inscriptions with and without occupations, just to see the difference of epigraphically active people vs epigraphically active and occupationaly proud people)


status_people<- LIST_all_people_harris %>% 
  select(status, occups_cats) %>% 
  filter(!is.na(status)) %>% 
  mutate(status_clean = str_replace_all(string = status, pattern = "\\?", replacement = "")) %>%
  count(status_clean, sort = TRUE) %>% 
    mutate(status_ratio = n/(nrow(LIST_all_people_harris)/100))
  
#save the results
write.csv(status_people, "LIST_status_people.csv",row.names = TRUE)

```

### Status by occupational categories



```{r}
# representation of status in individual occupational categories

LIST_all_people_harris %>% 
  select(`LIST-ID`, occups, occups_cats, status) %>% 
  filter(!is.na(occups)) %>% 
  filter(occups_cats == "Water-Management") %>% 
  count(status, sort = T) %>% 
  mutate(ratio = n/(sum(n)/100))
```




We have information about status from 98,376 only about 28,128, so roughly only one-third 

```{r}
# available status from all available people 
nrow(LIST_all_people_harris) / nrow(LIST_all_people_harris %>% filter(!is.na(status)))

# people with status
people_status <- LIST_all_people_harris %>% 
  select(status, occups_cats) %>% 
  filter(!is.na(status))

people_status %>% 
  count(occups_cats, sort = T) %>% 
  mutate(ratio = round(n/(sum(n)/100),2))

```

```{r}
people_status_clean<- people_status %>% 
  mutate(status_clean = str_replace_all(string = status, pattern = "\\?", replacement = "")) %>% 
  count(status_clean, sort = T) %>% 
  mutate(ratio = round(n/(sum(n)/100),2)) 

```

```{r}
#save the results
write.csv(people_status_clean, "LIST_status_people_clean.csv",row.names = TRUE)

```

```{r}
people_status %>% 
  filter(!is.na(occups_cats)) %>% 
  mutate(status_clean = str_replace_all(string = status, pattern = "\\?", replacement = "")) %>%
  count(status_clean, sort = T) %>% 
  mutate(ratio = round(n/(sum(n)/100),2))
```

