---
title: "Occupations in Greek inscriptions"
author: 
- Petra Hermankova^[Aarhus University, Denmark, https://orcid.org/0000-0002-6349-0540]
date: "`r format(Sys.Date())`"
output:
  html_document:
    theme: united
    toc: yes
    toc_float: true
    number_sections: true
    toc_depth: 3
    df_print: paged
---

# Initial setup

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(sf)
library(sfarrow)
library(tidyverse)
library(jsonlite)
library(tidyr)
library(ggplot2)
library(viridis)
library(arrow)
library(stringr)
```

```{r loading data}
GIST <- read_parquet("../../../data/large_data/GIST_v1-1.parquet")

occups <- read_tsv("../../../data/Occupations_GREEK.tsv")

```

# Detecting occupations

## Lemmatized text
```{r,finding occupations function}
# Helper function to escape special characters in regex patterns
escape_special_chars <- function(occupation) {
  # Escape regex special characters
  return(str_replace_all(occupation, "([\\^\\$\\.\\*\\+\\?\\(\\)\\[\\]\\{\\}\\|])", "\\\\\\1"))
}

# Function to find occupations with word boundaries and escaped characters
find_occupations <- function(text, occupations) {
  # Escape special characters and add word boundaries for each occupation
  occupations_escaped <- sapply(occupations, function(occ) paste0("\\b", escape_special_chars(occ), "\\b"))
  
  # Check for matches of each escaped occupation within the text
  matches <- occupations[sapply(occupations_escaped, function(occ_pattern) str_detect(text, occ_pattern))]
  
  # Return unique matches or NA if none found
  if (length(matches) == 0) {
    return(NA)
  } else {
    return(unique(matches))
  }
}


```

```{r, testing the methods on 1000 inscriptions}

GISTsample <- head(GIST, 1000)

GISTsample$occups <- lapply(GISTsample$lemmata, find_occupations, occupations = occups$Term)

GISTsample %>%
  unnest(occups) %>%
  count(occups, sort=T)

GISTsample %>% 
  filter(!is.na(occups)) -> eval_sample

```

```{r, detection methods evaluation 1 by 1}
eval_sample %>% 
  filter(PHI_ID == 1350) -> inscription 
pattern <- "ποιμήν"
  
sapply(inscription$lemmata, function(lemmata) {
  any(sapply(lemmata, function(pattern) str_detect(inscription, pattern)))
})


```

```{r, detection methods evaluation manual OUT}
eval_sample %>% 
  unnest(occups) %>% 
  select(PHI_ID, reference, occups)  -> manual_eval

write_csv(manual_eval, "manual_eval_detection_method_1000_OUT.csv")

```


Out of 1000 inscriptions, 29 had at least one occupation detected, 33 instances of occupations. More than 50% of the detected occupations is wrongly assigned. 

```{r, detection methods evaluation manual IN}
eval_done<- read_csv("manual_eval_detection_method_1000_IN.csv")

eval_done %>% 
  count(correct) %>% 
  mutate(ratio = n/(nrow(eval_done)/100))
```


```{r, full corpus detection}
# takes few hours to run, 15:56 - 

# Apply the detection function to each row in the 'text' column
GIST$occups <- lapply(GIST$lemmata, find_occupations, occupations = occups$Term)

GIST %>%
  unnest(occups) %>%
  count(occups, sort=T)

```

```{r, exploring the results in the entire corpus}
GIST %>% 
  filter(!is.na(occups)) %>% 
  unnest(occups) %>% 
  count(occups, sort=T)
# 4624 > 4655 inscriptions with at least one occupation
# 5389 > 5424 occurrences of occupation
# 151 > 158 unique occupations
```

Control for problematic lemmata
Inhabitatnt of Halai deme Ἁλαιεὺς vs fisherman ἁλιεύς
 
```{r, exploring problematic lemmata 1}
GIST %>% 
  #filter(occups %in% "ἁλιεύς") 
  filter(clean_text_interpretive_word %in% str_subset(clean_text_interpretive_word, "Ἁλαιεὺς|Ἁλαιεύς")) 
GIST %>% 
  filter(clean_text_interpretive_word %in% str_subset(clean_text_interpretive_word, "ἁλιεύς"))

```

```{r, fixing problematic lemmata ad-hoc}
GISTfix <-GIST %>% 
  filter(!clean_text_interpretive_word %in% str_subset(clean_text_interpretive_word, "Ἁλαιεὺς|Ἁλαιεύς"))
```

Apllying the fixes, there is still very high mistake rate.

```{r, results after problematic lemmata fixed}
GISTfix %>% 
  filter(!is.na(occups)) %>% 
  unnest(occups) %>% 
  count(occups, sort=T)
# 4624 > 4655 inscriptions with at least one occupation > after fix 4569
# 5389 > 5424 occurrences of occupation > 5331
# 151 > 158 unique occupations > 158
```


## Morphological dictionary
Old approach using the full declension from https://en.wiktionary.org/wiki/%E1%BC%81%CE%BB%CE%B9%CE%B5%CF%8D%CF%82
https://www.perseus.tufts.edu/hopper/text?doc=Perseus:text:1999.04.0057:entry=a(lieu/s
Fisherman, seaman, sailor

```{r, testing one occupation}

test<- c("ἁλιεύς", "ἁλιέως", "ἁλιῶς", "ἁλιεῖ", "ἁλιᾶ", "ἁλιέᾱ", "ἁλιεῦ",
         "ἁλιῆ", "ἁλιέοιν", 
        "ἁλιῆς", "ἁλιεῖς", "ἁλιέων", "ἁλιῶν", "ἁλιεῦσῐ","ἁλιεῦσῐν", "ἁλιέᾱς", "ἁλιᾶς")

GIST %>% 
   filter(clean_text_interpretive_word %in% str_subset(clean_text_interpretive_word, pattern = "ἁλιεύς"))

```
```{r, detecting all instances of one occupation in the full corpus}
# Combine words into a pattern, ensuring it’s not preceded by an uppercase Greek letter
# This version will use a lookbehind for spaces or word boundaries before each term
pattern <- paste0("(?<![[:upper:]])\\b(", paste(test, collapse = "|"), ")\\b")

# Add a new column with the detected term if found, otherwise NA
GIST <- GIST %>%
  mutate(detected_term = str_extract(clean_text_interpretive_word, pattern))

# Filter to show only rows with a detected term (optional)
filtered_data <- GIST %>%
  filter(!is.na(detected_term))

# Display the result
filtered_data

filtered_data$occups
```
Results:  Occurences from 485 to 5 !!!

### Full morphological form search
```{r, expanding the detection to all occupations and the full corpus}
library(dplyr)
library(stringr)

# takes few hours to run!

morphlist<- str_split(occups$Morphological_dictionary, pattern = ", ")

# Step 2: Define a function to check for individual term matches in `clean_text_interpretive_word`
find_individual_matches <- function(text, terms) {
  # For each term, check if it is present in `text` and store matches
  matches <- unlist(lapply(terms, function(term) {
    if (str_detect(text, fixed(term))) term else NULL  # fixed() for exact match
  }))
  unique(matches)  # Return unique matches
}

# Step 3: Apply the function row-wise on `GISTsample`
GIST_result <- GIST %>%
  rowwise() %>%
  mutate(
    # Apply function to find individual matches for each `clean_text_interpretive_word`
    occup_list_morph = list(find_individual_matches(clean_text_interpretive_word, unlist(morphlist)))
  ) %>%
  ungroup() %>%
  # Keep only rows with at least one match
  filter(lengths(occup_list_morph) > 0)

# Display results
GIST_result %>% 
  unnest(occup_list_morph) %>% 
  count(occup_list_morph, sort=T)
```

