---
title: "eci_HISCO"
output: html_document
date: "2026-01-17"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction
This markdown extends economic_complexity.Rmd, recomputing its results using HISCO classifications of the occupational words instead of occupational words directly. 

Note: run after economic_complexity.Rmd

## HISCO data and conversion functions
First, we read the HISCO dataset
```{r read updated dataset with HISCO data}
occupations_all_HISCO <- read.csv("../../data/economic_complexity/Occupations_Ancient_Mediterranean_LATIN_HISCO.csv")
```

We assign a HISCO level 4 data to standardised occupations
```{r assign HISCO level 4 data to each standardised occupation}
# RK: we know they are in the same order
occupations_HISCO <- occupations_all
occupations_HISCO$HISCO <- occupations_all_HISCO$HISCO_codes
occupations_HISCO$HISCO4 <- str_sub(occupations_HISCO$HISCO,1,4)

# extract info on occupations' HISCO category
occupations_HISCO <- unique(occupations_HISCO[,c("std_spelling","HISCO4")])

# data issue: some professions have two different HISCO codes: we assign only the first one
HISCO_issues <-  occupations_HISCO$std_spelling[which(duplicated(occupations_HISCO$std_spelling))]
for (i in HISCO_issues){
  occupations_HISCO$HISCO4[occupations_HISCO$std_spelling ==i] <-
    occupations_HISCO$HISCO4[occupations_HISCO$std_spelling ==i][1]
}
occupations_HISCO <- unique(occupations_HISCO[,c("std_spelling","HISCO4")])

# initialise new column of HISCOs
inscriptions_w_occupations_in_provinces$occup_HISCO <- rep(list(NA),nrow(inscriptions_w_occupations_in_provinces))

# assign HISCO to the occupations
for (i in seq(nrow(inscriptions_w_occupations_in_provinces))){
  # extract an inscription
  inscription <- inscriptions_w_occupations_in_provinces[i,]
  # find the occupations in the inscription and associate it with the HISCO
  inscription_occ <- data.frame(std_spelling = unlist(inscription$occups_std)) %>%
    left_join(occupations_HISCO,
              by = "std_spelling")
  # save the standard spellings for the occupations in the inscription
  inscriptions_w_occupations_in_provinces$occup_HISCO[[i]] <- inscription_occ$HISCO4
}

# same for countries
inscriptions_w_occupations_in_countries$occup_HISCO <- rep(list(NA),nrow(inscriptions_w_occupations_in_countries))

# assign HISOC to the occupations
for (i in seq(nrow(inscriptions_w_occupations_in_countries))){
  # extract an inscription
  inscription <- inscriptions_w_occupations_in_countries[i,]
  # find the occupations in the inscription and associate it with the HISCO
  inscription_occ <- data.frame(std_spelling = unlist(inscription$occups_std)) %>%
    left_join(occupations_HISCO,
              by = "std_spelling")
  # save the standard spellings for the occupations in the inscription
  inscriptions_w_occupations_in_countries$occup_HISCO[[i]] <- inscription_occ$HISCO4
}
```

We adapt the function to construct occupation tables so it works with HISCO data
```{r modified table of occupation function for HISCO}
table_of_occupations_HISCO <- function(insc_w_occ){
  # ensure that the info on geographical unit (gu) is included in the input
  gu_type <- which_gu(insc_w_occ)

  # create list of geographical units
  if (gu_type == "province") {
    gu <- provinces$province
  } else if (gu_type == "country") {
    # note that some disputed areas in the intersections are not assigned to any country, hence exclude na
    gu <- na.omit(unique(intersections$country))
  } # no need to check for other cases as which_gu already takes care of it
  
  # unnest the list of occupations in each inscription (based on the standard occupational terms)
  insc_w_occ_unnested <- unnest(insc_w_occ,"occup_HISCO")
  
  # initialise table
  occ_table <- table(insc_w_occ_unnested[[gu_type]],
                     insc_w_occ_unnested$occup_HISCO)

  # add zero rows for gus without inscriptions
  gu_not_in_occ_table <- gu[!gu%in%insc_w_occ_unnested[[gu_type]]]
  occ_table_zeros <- matrix(0, nrow= length(gu_not_in_occ_table), ncol = ncol(occ_table))
  rownames(occ_table_zeros) <- gu_not_in_occ_table
  occ_table <- rbind(occ_table,occ_table_zeros) %>% as.table()

  # sort the table based on row and column sums
  occ_table <- sort_table(occ_table)
  
  # return occupation table
  return(occ_table)
} 
```

Finally, we compute the occupation tables based on HISCO level 4 codes.
```{r compute HISCO occupational tables}
province_table_HISCO <- table_of_occupations_HISCO(inscriptions_w_occupations_in_provinces)
country_table_HISCO <- table_of_occupations_HISCO(inscriptions_w_occupations_in_countries)
```

We plot and save the results
```{r plot and save heatmaps}
# define heatmaps
province_table_HISCO_heatmap <- occ_heat_map(province_table_HISCO)
country_table_HISCO_heatmap <- occ_heat_map(country_table_HISCO)

# plot and save
plot_and_save(province_table_HISCO_heatmap,width = 15, height = 10)
plot_and_save(country_table_HISCO_heatmap,width = 15, height = 10)
```

We also create a function to automatically convert table of occupations from standard spelling to HISCO codes
```{r function to convert a table of occupations into a HISCO table}
# define map from occupation standard spelling to HISCO category
occ_to_HISCO <- setNames(occupations_HISCO$HISCO4, occupations_HISCO$std_spelling)

# function to convert occupation table to HISCO table
table_to_HISCO <- function(tbl){
  df_HISCO <- sapply(
    split(colnames(tbl), occ_to_HISCO[colnames(tbl)]),
    function(cols) rowSums(tbl[, cols, drop = FALSE])
    )
  tbl_HISCO <- sort_table(as.table(df_HISCO))
  return(tbl_HISCO)
}

# check if correct
# example_table <- province_table_greek_latin_lists$r_0$it_1
# table_to_HISCO(example_table)
```

## Analysis
### Effects of individual biases on eci
Adapt the function to assess the impact of individual bias corrections
```{r modify individual_bias_eci_stats for HISCO}
# this function computes eci stats after each bias correction procedure
individual_bias_eci_stats_HISCO <- function(original_tb){
  # identify gu type
  gu_type <- which_gu(original_tb)
  
  # initialise stats
  gu_stats <- list()
  
  # ensure that the occupation table has no zero rows
  original_tb <- original_tb[rowSums(original_tb)>0,]
  
  # convert to HISCO
  original_tb_HISCO <- table_to_HISCO(original_tb)
  
  # compute values for the original HISCO table
  gu_stats$original <- economic_complexity(original_tb_HISCO)
  
  # language bias
  language_stats <- list()
  for (m in paste0("r_",max_distances[1:2])){
    gu_table_greek_latin_lists <-
      get(
        paste0(
          gu_type,
          "_table_greek_latin_lists"
          )
        )[[m]][1:100]
    gu_table_HISCO_greek_latin_lists <- 
      lapply(
        gu_table_greek_latin_lists, 
        table_to_HISCO)
    
    # compute eci for this list
    ecis_m <- eci_list(gu_table_HISCO_greek_latin_lists)
    language_stats[[m]] <- eci_stats(ecis_m,gu_type)
    }
  gu_stats[["language"]] <- language_stats
  print("language done")
  
  # status bias
  status_tbs <- status_bias_correction_stochastic(original_tb,15,30,n_samples=100)
  status_tbs_HISCO <- lapply(status_tbs, table_to_HISCO)
  status_ecis <- eci_list(status_tbs_HISCO)
  gu_stats[["status"]] <- eci_stats(status_ecis,gu_type)
  print("status done")

  # gender bias
  gender_tbs <- lapply(gender_ratio_v[1:100], 
                       function(gender_ratio) gender_bias(original_tb, gender_ratio))
  gender_tbs_HISCO <- lapply(gender_tbs, table_to_HISCO)
  gender_ecis <- eci_list(gender_tbs_HISCO)
  gu_stats[["gender"]] <- eci_stats(gender_ecis,gu_type)
  print("gender done")

  # agricola
  agricola_tbs <- lapply(seq(0,1,by=0.01),
                         function(s) agricola_correction(original_tb,p=0.95,s))
  agricola_tbs_HISCO <- lapply(agricola_tbs, table_to_HISCO)
  agricola_ecis <- eci_list(agricola_tbs_HISCO)
  gu_stats[["agricola"]] <- eci_stats(agricola_ecis,gu_type)
  print("agricola done")

  # research
  research_tbs <- lapply(seq(0,1,by=0.01),
                         function(p) research_correction(original_tb,p))
  research_tbs_HISCO <- lapply(research_tbs, table_to_HISCO)
  research_ecis <- eci_list(research_tbs_HISCO)
  gu_stats[["research"]] <- eci_stats(research_ecis,gu_type)
  print("research done")

  return(gu_stats)
}
```

Apply the function and study the effects on all provinces
```{r compute the stats for all provinces and countries eci}
province_individual_bias_eci_stats_HISCO <- individual_bias_eci_stats_HISCO(province_table)
# save the results
saveRDS(province_individual_bias_eci_stats_HISCO, file = "../../results/economic_complexity/province_individual_bias_eci_stats_HISCO.rds")

country_individual_bias_eci_stats_HISCO <- individual_bias_eci_stats_HISCO(country_table)
# save the results
saveRDS(country_individual_bias_eci_stats_HISCO, file = "../../results/economic_complexity/country_individual_bias_eci_stats_HISCO.rds")
```

We then compute the stats for filtered provinces and countries (conservative approach)
```{r compute the stats for filtered provinces and countries eci}
# provinces
filtered_province_individual_bias_eci_stats_HISCO <-
  individual_bias_eci_stats_HISCO(filtered_province_table)
# save the results
saveRDS(filtered_province_individual_bias_eci_stats_HISCO, file = "../../results/economic_complexity/filtered_province_individual_bias_eci_stats_HISCO.rds")

# countries
filtered_country_individual_bias_eci_stats_HISCO <- individual_bias_eci_stats_HISCO(filtered_country_table)
# save the results
saveRDS(filtered_country_individual_bias_eci_stats_HISCO, file = "../../results/economic_complexity/filtered_country_individual_bias_eci_stats_HISCO.rds")
```

We plot these as boxplots
```{r boxplot of effects of individual biases on eci}
for (exclude_Rome_gu in c(TRUE,FALSE)){
  for (gu in c("province","country")) {
  bias_tb <- get(paste0(gu,"_individual_bias_eci_stats_HISCO"))
  #bias_tb <- get(paste0("filtered_",gu,"_individual_bias_eci_stats"))

  for (bias in c("language","status","gender","agricola","research")) {
    tb <- bias_tb[[bias]]
    if (bias == "language"){
      for (m in paste0("r_",max_distances[1:2])) {
        # extract eci
        eci_m <- tb[[m]]
        original_eci <- bias_tb$original$eci
        
        # select eci values 
        eci_values <- eci_m[,c(gu,paste0("eci_",1:100))]
        if (exclude_Rome_gu){
          eci_values <- eci_values[!eci_values[[gu]] %in% 
                                     c("ITA","Latium et Campania (Regio I)"),]
          original_eci <- original_eci[!original_eci[[gu]] %in% 
                                         c("ITA","Latium et Campania (Regio I)"),]
          }
        
        # tranform into a long df for ggplot boxplot
        eci_long <- eci_values %>%
          pivot_longer(cols = -gu, names_to = "sample", values_to = "eci")
        
        # sort the gu names in ascending order of median eci  
        gu_order <- eci_m[[gu]][order(eci_m$median)]
        eci_long[[gu]] <- factor(eci_long[[gu]], levels = gu_order)
        
        # define title
        boxplot_title = paste0(
          paste0("Eci by ", gu," based on ", bias," bias correction ",m),
          ifelse(exclude_Rome_gu,", excluding Rome's region","")) 
        
        # define boxplot
        eci_boxplot <- ggplot(eci_long, aes(x = !!sym(gu), y = eci)) +
          geom_boxplot(fill= "black", alpha=0.2) +
          geom_point(data = original_eci, 
                     aes(x = !!sym(gu), y = eci, color = "original eci"), size = 2) +
          scale_color_manual(values = "red", 
                     name = "Legend", 
                     labels = "original eci") +
          theme_minimal() +
          labs(title = boxplot_title, x = sym(gu), y = "eci") +
          theme(axis.text.x = element_text(angle = 45, hjust = 1))
        
        # save
        boxplot_name <- paste0(
          paste0(gu,"_",bias,"_bias_correction_eci_boxplot_",m,"_HISCO"),
          ifelse(exclude_Rome_gu,"_excluding_Rome_s_region","")) 
        # boxplot_name <- paste0(
        #  paste0("filtered_",gu,"_",bias,"_bias_correction_eci_boxplot_",m),
        #  ifelse(exclude_Rome_gu,"_excluding_Rome_s_region","")) 
        plot_and_save(eci_boxplot,
                      title = boxplot_name,
                      height = 5,
                      width = 10)
      } 
    } else {
        # extract eci values 
        eci_values <- tb[,c(gu,paste0("eci_",1:100))]
        original_eci <- bias_tb$original$eci
        
        #select eci values
         if (exclude_Rome_gu){
           eci_values <- eci_values[!eci_values[[gu]] %in% 
                                      c("ITA","Latium et Campania (Regio I)"),]
           original_eci <- original_eci[!original_eci[[gu]] %in% 
                                          c("ITA","Latium et Campania (Regio I)"),]
           }
      
        # tranform into a long df for ggplot boxplot
        eci_long <- eci_values %>%
          pivot_longer(cols = -gu, names_to = "sample", values_to = "eci")
        
        # sort the gu names in ascending order of median eci  
        gu_order <- tb[[gu]][order(tb$median)]
        eci_long[[gu]] <- factor(eci_long[[gu]], levels = gu_order)
        
        # define title
         boxplot_title = paste0(
          paste0("Eci by ", gu," based on ", bias," bias correction"),
          ifelse(exclude_Rome_gu,", excluding Rome's region","")) 
        
        # define boxplot
        eci_boxplot <- ggplot(eci_long, aes(x = !!sym(gu), y = eci)) +
          geom_boxplot(fill= "black", alpha=0.2) +
          geom_point(data = original_eci, 
                     aes(x = !!sym(gu), y = eci, color = "original eci"), size = 2) +
          scale_color_manual(values = "red", 
                     name = "Legend", 
                     labels = "original eci") +
          theme_minimal() +
          labs(title = boxplot_title, x = sym(gu), y = "eci") +
          theme(axis.text.x = element_text(angle = 45, hjust = 1))
        
        # save
        boxplot_name <- paste0(
          paste0(gu,"_",bias,"_bias_correction_eci_boxplot_HISCO"),
          ifelse(exclude_Rome_gu,"_excluding_Rome_s_region","")) 
        #boxplot_name <- paste0(
        # paste0("filtered_",gu,"_",bias,"_bias_correction_eci_boxplot"),
        #  ifelse(exclude_Rome_gu,"_excluding_Rome_s_region","")) 
        plot_and_save(eci_boxplot,
                      title = boxplot_name,
                      height = 5,
                      width = 10)
    }
    print(bias)
  }
  print(gu)
}
}
```

### Combined effects of different bias corrections on eci
We now adapt the function combining all the required correction strategies
```{r eci stats after all bias corrections}
# this function computes eci stats after all bias correction procedures have been applied
all_bias_eci_stats_HISCO <- function(original_tb,
                               language_it = 10, # cases considered
                               status_it = 2, 
                               gender_it = 2,
                               agricola_it = 5,
                               research_it = 5){
  # identify gu type
  gu_type <- which_gu(original_tb)
  
  # initialise stats
  gu_stats <- list()
  
  # ensure that the occupation table has no zero rows
  original_tb <- original_tb[rowSums(original_tb)>0,]
  
  # convert to HISCO
  original_tb_HISCO <- table_to_HISCO(original_tb)
  
  # compute values for the original table
  gu_stats$original <- economic_complexity(original_tb_HISCO)
  
  # loop and add
  for (m in paste0("r_",max_distances[c(1,2)])){
    languages_tbl_m <- get(paste0(gu_type,"_table_greek_latin_lists"))[[m]][seq_len(language_it)]
    bias_type_m <- paste0("language_r_",m)
    
    for (l_it in seq_len(language_it)) {
      tb_l_it <- languages_tbl_m[[l_it]]
      status_tbs <- status_bias_correction_stochastic(tb_l_it,15,30,n_samples=status_it)
      bias_type_l <- paste0(bias_type_m,"_l_it_",l_it,"_status")
      
      for (s_it in seq_len(status_it)) {
        tb_s_it <- status_tbs[[s_it]]
        gender_tbs <- lapply(gender_ratio_v, 
                       function(gender_ratio) gender_bias(tb_s_it, gender_ratio))
        bias_type_s <- paste0(bias_type_l,"_s_it_",s_it,"_gender")

        for (g_it in seq_len(gender_it)) {
          tb_g_it <- gender_tbs[[g_it]]
          agricola_tbs <- lapply(seq(0,1,length.out=agricola_it),
                         function(s) agricola_correction(tb_g_it,p=0.95,s))
          bias_type_g <- paste0(bias_type_s,"_g_it_",g_it,"_agricola")

          for (a_it in seq_len(agricola_it)) {
            tb_a_it <- agricola_tbs[[a_it]]
            research_tbs <- lapply(seq(0,1,length.out=research_it),
                                   function(p) research_correction(tb_a_it,p))
            bias_type_a <- paste0(bias_type_g,"_a_it_",a_it,"_research")
            
            # convert to HISCO
            research_tbs_HISCO <- lapply(research_tbs, table_to_HISCO)
            research_ecis <- eci_list(research_tbs_HISCO)
            gu_stats[[bias_type_a]] <- eci_stats(research_ecis,gu_type)
            print(bias_type_a)
          }
        } 
      }
    }
  }
  return(gu_stats)
}

# this function does the same, without language bias
all_but_language_bias_eci_stats_HISCO <- function(original_tb,
                                            status_it = 2, # cases considered
                                            gender_it = 2,
                                            agricola_it = 5,
                                            research_it = 5){
  # identify gu type
  gu_type <- which_gu(original_tb)
  
  # initialise stats
  gu_stats <- list()
  
  # ensure that the occupation table has no zero rows
  original_tb <- original_tb[rowSums(original_tb)>0,]
  
  # convert to HISCO
  original_tb_HISCO <- table_to_HISCO(original_tb)
  
  # compute values for the original table
  gu_stats$original <- economic_complexity(original_tb_HISCO)

  # apply status correction
  status_tbs <- status_bias_correction_stochastic(original_tb,15,30,n_samples=status_it)

  # loop and add
  for (s_it in seq_len(status_it)) {
    tb_s_it <- status_tbs[[s_it]]
    gender_tbs <- lapply(gender_ratio_v, 
                   function(gender_ratio) gender_bias(tb_s_it, gender_ratio))
    bias_type_s <- paste0("status_s_it_",s_it,"_gender")

    for (g_it in seq_len(gender_it)) {
      tb_g_it <- gender_tbs[[g_it]]
      agricola_tbs <- lapply(seq(0,1,length.out=agricola_it),
                     function(s) agricola_correction(tb_g_it,p=0.95,s))
      bias_type_g <- paste0(bias_type_s,"_g_it_",g_it,"_agricola")

      for (a_it in seq_len(agricola_it)) {
        tb_a_it <- agricola_tbs[[a_it]]
        research_tbs <- lapply(seq(0,1,length.out=research_it),
                               function(p) research_correction(tb_a_it,p))
        bias_type_a <- paste0(bias_type_g,"_a_it_",a_it,"_research")
             
        # convert to HISCO
        research_tbs_HISCO <- lapply(research_tbs, table_to_HISCO)
        research_ecis <- eci_list(research_tbs_HISCO)
        gu_stats[[bias_type_a]] <- eci_stats(research_ecis,gu_type)
        print(bias_type_a)
      }
    } 
  }
  return(gu_stats)
}
```

Apply all corrections
```{r compute the stats for provinces and countries eci after all bias corrections}
# provinces
province_all_bias_eci_stats_HISCO <- all_bias_eci_stats_HISCO(province_table)
# save the results
saveRDS(province_all_bias_eci_stats_HISCO, file = "../../results/economic_complexity/province_all_bias_eci_stats_HISCO.rds")

# countries
country_all_bias_eci_stats_HISCO <- all_bias_eci_stats_HISCO(country_table)
# save the results
saveRDS(country_all_bias_eci_stats_HISCO, file = "../../results/economic_complexity/country_all_bias_eci_stats_HISCO.rds")
```
 
Apply all corrections but language to the filtered tables (conservative approach)
```{r compute the stats for filtered provinces and countries eci after all but language bias corrections}
# provinces
filtered_province_all_but_language_bias_eci_stats_HISCO <- all_but_language_bias_eci_stats_HISCO(filtered_province_table)
# save the results
saveRDS(filtered_province_all_but_language_bias_eci_stats_HISCO, file = "../../results/economic_complexity/filtered_province_all_but_language_bias_eci_stats_HISCO.rds")

# countries
filtered_country_all_but_language_bias_eci_stats_HISCO <- all_but_language_bias_eci_stats_HISCO(filtered_country_table)
# save the results
saveRDS(filtered_country_all_but_language_bias_eci_stats_HISCO, file = "../../results/economic_complexity/filtered_country_all_but_language_bias_eci_stats_HISCO.rds")
```

### Alternative functions for eci (df as output)
Alternative corrective functions
```{r eci stats after all bias corrections}
# this function computes eci stats after all bias correction procedures have been applied
all_bias_eci_HISCO_df <- function(original_tb,
                            language_it = 10, # cases considered
                            status_it = 2, 
                            gender_it = 2,
                            agricola_it = 5,
                            research_it = 5){
  # identify gu type
  gu_type <- which_gu(original_tb)
  
  # ensure that the occupation table has no zero rows
  original_tb <- original_tb[rowSums(original_tb)>0,]
  
  #retrieve language tables
  languages_tbls <- get(paste0(gu_type,"_table_greek_latin_lists"))
  
  # initialise bias dataframe
  case_r_0 <- languages_tbls$r_0$it_1
  case_r_0.001 <- languages_tbls$r_0.001$it_1
  nrows_case_r_0 <- nrow(case_r_0[rowSums(case_r_0)>0,])
  nrows_case_r_0.001 <- nrow(case_r_0.001[rowSums(case_r_0.001)>0,])

  nrows <- language_it*status_it*gender_it*agricola_it*research_it*(nrows_case_r_0+nrows_case_r_0.001)
  bias_df <- data.frame(matrix(NA, ncol = 8, nrow = nrows)) %>% 
    setNames(c(gu_type,"m","l_it","s_it","g_it","a_it","r_it","eci"))
  
  # define indices for filling the df
  index_start <- 1
    
  # loop and add
  for (m in paste0("r_",max_distances[c(1,2)])){
    languages_tbl_m <- languages_tbls[[m]][seq_len(language_it)]
    nrows_case <- get(paste0("nrows_case_",m))
    
    for (l_it in seq_len(language_it)) {
      tb_l_it <- languages_tbl_m[[l_it]]
      status_tbs <- status_bias_correction_stochastic(tb_l_it,15,30,n_samples=status_it)

      for (s_it in seq_len(status_it)) {
        tb_s_it <- status_tbs[[s_it]]
        gender_tbs <- lapply(gender_ratio_v, 
                       function(gender_ratio) gender_bias(tb_s_it, gender_ratio))

        for (g_it in seq_len(gender_it)) {
          tb_g_it <- gender_tbs[[g_it]]
          agricola_tbs <- lapply(seq(0,1,length.out=agricola_it),
                         function(s) agricola_correction(tb_g_it,p=0.95,s))

          for (a_it in seq_len(agricola_it)) {
            tb_a_it <- agricola_tbs[[a_it]]
            research_tbs <- lapply(seq(0,1,length.out=research_it),
                                   function(p) research_correction(tb_a_it,p))

            # convert to HISCO
            research_tbs_HISCO <- lapply(research_tbs, table_to_HISCO)
            research_ecis <- eci_list(research_tbs_HISCO)
            
            for (r_it in seq(research_it)){
              eci_r_it <- research_ecis[[r_it]]
              
              # define end index
              index_end <- index_start+nrows_case-1
              
              # update dataframes 
              bias_df[index_start:index_end,c(gu_type,"eci")] <- eci_r_it
              bias_df[index_start:index_end,c("m","l_it","s_it","g_it","a_it","r_it")]<-
                rep(c(m,l_it,s_it,g_it,a_it,r_it), each = nrows_case)
              
              # update start index
              # find indices
              index_start <- index_end +1
              
              # check progress
              # print(paste0(index_end," out of ",nrows))
            }
          }
        } 
      }
    }
  }
  return(bias_df)
}

# this function does the same, without language bias 
all_but_language_bias_eci_HISCO_df <- function(original_tb,
                                            status_it = 2, # cases considered
                                            gender_it = 2,
                                            agricola_it = 5,
                                            research_it = 5){
  # identify gu type
  gu_type <- which_gu(original_tb)
  
  # ensure that the occupation table has no zero rows
  original_tb <- original_tb[rowSums(original_tb)>0,]
  
  # n_rows
  nrow_case <- nrow(original_tb)
  nrows <- status_it*gender_it*agricola_it*research_it*nrow_case
  
  # initialise result df
  bias_df <- data.frame(matrix(NA, ncol = 6, nrow = nrows)) %>% 
    setNames(c(gu_type,"s_it","g_it","a_it","r_it","eci"))
  
  # define indices for filling the df
  index_start <- 1
  
  # apply status correction
  status_tbs <- status_bias_correction_stochastic(original_tb,15,30,n_samples=status_it)

  # loop and add
  for (s_it in seq_len(status_it)) {
    tb_s_it <- status_tbs[[s_it]]
    gender_tbs <- lapply(gender_ratio_v, 
                   function(gender_ratio) gender_bias(tb_s_it, gender_ratio))

    for (g_it in seq_len(gender_it)) {
      tb_g_it <- gender_tbs[[g_it]]
      agricola_tbs <- lapply(seq(0,1,length.out=agricola_it),
                     function(s) agricola_correction(tb_g_it,p=0.95,s))

      for (a_it in seq_len(agricola_it)) {
        tb_a_it <- agricola_tbs[[a_it]]
        research_tbs <- lapply(seq(0,1,length.out=research_it),
                               function(p) research_correction(tb_a_it,p))
        
          # convert to HISCO
          research_tbs_HISCO <- lapply(research_tbs, table_to_HISCO)
          research_ecis <- eci_list(research_tbs_HISCO)

         for (r_it in seq(research_it)){
              eci_r_it <- research_ecis[[r_it]]
              
              # define end index
              index_end <- index_start+nrow_case-1
              
              # update dataframes 
              bias_df[index_start:index_end,c(gu_type,"eci")] <- eci_r_it
              bias_df[index_start:index_end,c("s_it","g_it","a_it","r_it")]<-
                rep(c(s_it,g_it,a_it,r_it), each = nrow_case)
              
              # update start index
              # find indices
              index_start <- index_end +1
              
              print(paste0(index_end," out of ",nrows))
            }
        
      }
    } 
  }
  return(bias_df)
}
```

Apply alternative corrections to the original tables
```{r compute the eci for provinces and countries after all bias corrections}
# provinces
province_all_bias_eci_HISCO_df <- all_bias_eci_HISCO_df(province_table)
# save the results
write.csv(province_all_bias_eci_HISCO_df, file = "../../results/economic_complexity/province_all_bias_eci_df_HISCO.csv")

# countries
country_all_bias_eci_HISCO_df <- all_bias_eci_HISCO_df(country_table)
# save the results
write.csv(country_all_bias_eci_HISCO_df, file = "../../results/economic_complexity/country_all_bias_eci_df_HISCO.csv")
```

Apply alternative corrections but language to the filtered tables (conservative approach)
```{r compute the eci for filtered provinces and countries after all but language bias corrections}
# provinces
filtered_province_all_but_language_bias_eci_HISCO_df <- all_but_language_bias_eci_HISCO_df(filtered_province_table)
# save the results
write.csv(filtered_province_all_but_language_bias_eci_HISCO_df, file = "../../results/economic_complexity/filtered_province_all_but_language_bias_eci_df_HISCO.csv")

# countries
filtered_country_all_but_language_bias_eci_HISCO_df <- all_but_language_bias_eci_HISCO_df(filtered_country_table)
# save the results
write.csv(filtered_country_all_but_language_bias_eci_HISCO_df, file = "../../results/economic_complexity/filtered_country_all_but_language_bias_eci_df_HISCO.csv")
```

### Alternative functions for occupation tables
Produce occupation tables instead
```{r functions for occupation tables all biases}
# this function computes tbls after all bias correction procedures have been applied
all_bias_tbs_HISCO <- function(original_tb,
                           language_it = 10, # cases considered
                            status_it = 2, 
                            gender_it = 2,
                            agricola_it = 5,
                            research_it = 5){
  # identify gu type
  gu_type <- which_gu(original_tb)
  
  # ensure that the occupation table has no zero rows
  original_tb <- original_tb[rowSums(original_tb)>0,]
  
  #retrieve language tables
  languages_tbls <- get(paste0(gu_type,"_table_greek_latin_lists"))
  
  # initialise tb list
  tbs_list <- list()
    
  # loop and add
  for (m in paste0("r_",max_distances[c(1,2)])){
    languages_tbl_m <- languages_tbls[[m]][seq_len(language_it)]

    for (l_it in seq_len(language_it)) {
      tb_l_it <- languages_tbl_m[[l_it]]
      status_tbs <- status_bias_correction_stochastic(tb_l_it,15,30,n_samples=status_it)

      for (s_it in seq_len(status_it)) {
        tb_s_it <- status_tbs[[s_it]]
        gender_tbs <- lapply(gender_ratio_v, 
                       function(gender_ratio) gender_bias(tb_s_it, gender_ratio))

        for (g_it in seq_len(gender_it)) {
          tb_g_it <- gender_tbs[[g_it]]
          agricola_tbs <- lapply(seq(0,1,length.out=agricola_it),
                         function(s) agricola_correction(tb_g_it,p=0.95,s))

          for (a_it in seq_len(agricola_it)) {
            tb_a_it <- agricola_tbs[[a_it]]
            research_tbs <- lapply(seq(0,1,length.out=research_it),
                                   function(p) research_correction(tb_a_it,p))
            
            # convert to HISCO
            research_tbs_HISCO <- lapply(research_tbs, table_to_HISCO)

            for (r_it in seq(research_it)){
              tb_name <- ifelse(m=="r_0","without_neighbours","with_neighbours")
              tb_name <- paste0(tb_name,
                                "_l_it_",l_it,
                                "_s_it_",s_it,
                                "_g_it_",g_it,
                                "_a_it_",a_it,
                                "_r_it_",r_it)
              
              tbs_list[[tb_name]] <- research_tbs_HISCO[[r_it]]
            
              # check progress
              #print(tb_name)
            }
          }
        } 
      }
    }
  }
  return(tbs_list)
}


# this function does the same, without language bias 
all_but_language_tbs_HISCO<- function(original_tb,
                                status_it = 2, # cases considered
                                gender_it = 2,
                                agricola_it = 5,
                                research_it = 5){
  # identify gu type
  gu_type <- which_gu(original_tb)
  
  # ensure that the occupation table has no zero rows
  original_tb <- original_tb[rowSums(original_tb)>0,]
  
  # initialise tb list
  tbs_list <- list()
  
  # apply status correction
  status_tbs <- status_bias_correction_stochastic(original_tb,15,30,n_samples=status_it)

  # loop and add
  for (s_it in seq_len(status_it)) {
    tb_s_it <- status_tbs[[s_it]]
    gender_tbs <- lapply(gender_ratio_v, 
                   function(gender_ratio) gender_bias(tb_s_it, gender_ratio))

    for (g_it in seq_len(gender_it)) {
      tb_g_it <- gender_tbs[[g_it]]
      agricola_tbs <- lapply(seq(0,1,length.out=agricola_it),
                     function(s) agricola_correction(tb_g_it,p=0.95,s))

      for (a_it in seq_len(agricola_it)) {
        tb_a_it <- agricola_tbs[[a_it]]
        research_tbs <- lapply(seq(0,1,length.out=research_it),
                               function(p) research_correction(tb_a_it,p))
        
        # convert to HISCO
        research_tbs_HISCO <- lapply(research_tbs, table_to_HISCO)
         
         for (r_it in seq(research_it)){
              tb_name <- paste0("s_it_",s_it,
                                "_g_it_",g_it,
                                "_a_it_",a_it,
                                "_r_it_",r_it)
              
              tbs_list[[tb_name]] <- research_tbs_HISCO[[r_it]]
            
              # check progress
              #print(tb_name)
              }
      }
    } 
  }
  return(tbs_list)
}
```

Apply alternative corrections to the original tables
```{r compute the tables for provinces and countries after all bias corrections}
# provinces
province_all_bias_tbs_HISCO <- all_bias_tbs_HISCO(province_table)
# save the results
saveRDS(province_all_bias_tbs_HISCO, file = "../../results/economic_complexity/province_all_bias_tbs_HISCO.rds")

# countries
country_all_bias_tbs_HISCO <- all_bias_tbs_HISCO(country_table)
# save the results
saveRDS(country_all_bias_tbs_HISCO, file = "../../results/economic_complexity/country_all_bias_tbs_HISCO.rds")
```

Apply alternative corrections but language to the filtered tables
```{r compute the tables for filtered provinces and countries after all but language bias corrections}
# provinces
filtered_province_all_but_language_bias_tbs_HISCO <- all_but_language_tbs_HISCO(filtered_province_table)
# save the results
saveRDS(filtered_province_all_but_language_bias_tbs_HISCO, file = "../../results/economic_complexity/filtered_province_all_but_language_bias_tbs_HISCO.rds")

# countries
filtered_country_all_but_language_bias_tbs_HISCO <- all_but_language_tbs_HISCO(filtered_country_table)
# save the results
saveRDS(filtered_country_all_but_language_bias_tbs_HISCO, file = "../../results/economic_complexity/filtered_country_all_but_language_bias_tbs_HISCO.rds")
```

### Exploratory visuals
Use the eci tables to create boxplots
```{r ecis boxplots}
# with all gu, and language bias (speculative)
for (exclude_Rome_gu in c(TRUE,FALSE)) {
  for (gu in c("province","country")) {
     for (m in paste0("r_",max_distances[1:2])) {
       # extract eci
       eci_m <- get(paste0(gu,"_all_bias_eci_HISCO_df"))
       eci_values <- eci_m[eci_m$m==m,c(gu,"eci")]
       original_eci <- get(paste0(gu,"_all_bias_eci_stats_HISCO"))$original$eci
        
       # specify plot
       boxplot_title <- paste0("Bias-corrected HISCO eci by ",gu)
       boxplot_title <- ifelse(m=="r_0",
                               paste0(boxplot_title,", language correction without neighbours"),
                               paste0(boxplot_title,", language correction with neighbours"))

       # select eci values 
       if (exclude_Rome_gu){
          eci_values <- eci_values[!eci_values[[gu]] %in% 
                                     c("ITA","Latium et Campania (Regio I)"),]
          original_eci <- original_eci[!original_eci[[gu]] %in% 
                                         c("ITA","Latium et Campania (Regio I)"),]
          boxplot_title <- paste0(boxplot_title, ", excluding Rome's region")
       }
       
       
      # define boxplot
      eci_boxplot <- ggplot(eci_values, 
                            aes(x = reorder(!!sym(gu), eci, FUN = median), y = eci)) +
        geom_boxplot(fill= "black", alpha=0.2) +
        geom_point(data = original_eci, 
                   aes(x = !!sym(gu), y = eci, color = "original eci"), size = 2) +
        scale_color_manual(values = "red", 
                   name = "Legend", 
                   labels = "original eci") +
        theme_minimal() +
        labs(title = boxplot_title, x = sym(gu), y = "eci") +
        theme(axis.text.x = element_text(angle = 45, hjust = 1))

          
      # save
      boxplot_name <- str_replace_all(boxplot_title," ","_")
      plot_and_save(eci_boxplot,
                    title = boxplot_name,
                    height = 5,
                    width = 10)
      
      # check progress
      print(boxplot_name)
    }
  }
}

# with filtered gu, and without language (conservative)
for (exclude_Rome_gu in c(TRUE,FALSE)) {
  for (gu in c("province","country")) {
    # extract eci
     eci_values <- get(paste0("filtered_",gu,"_all_but_language_bias_eci_HISCO_df"))
     eci_values <- eci_values[,c(gu,"eci")]
     original_eci <- get(paste0("filtered_",gu,"_all_but_language_bias_eci_stats_HISCO"))$original$eci
      
     # specify plot
     boxplot_title <- paste0("Bias-corrected HISCO eci by selected ",gu)

     # select eci values 
     if (exclude_Rome_gu){
        eci_values <- eci_values[!eci_values[[gu]] %in% 
                                   c("ITA","Latium et Campania (Regio I)"),]
        original_eci <- original_eci[!original_eci[[gu]] %in% 
                                       c("ITA","Latium et Campania (Regio I)"),]
        boxplot_title <- paste0(boxplot_title, ", excluding Rome's region")
     }
     
     
    # define boxplot
    eci_boxplot <- ggplot(eci_values, 
                          aes(x = reorder(!!sym(gu), eci, FUN = median), y = eci)) +
      geom_boxplot(fill= "black", alpha=0.2) +
      geom_point(data = original_eci, 
                 aes(x = !!sym(gu), y = eci, color = "original eci"), size = 2) +
      scale_color_manual(values = "red", 
                 name = "Legend", 
                 labels = "original eci") +
      theme_minimal() +
      labs(title = boxplot_title, x = sym(gu), y = "eci") +
      theme(axis.text.x = element_text(angle = 45, hjust = 1))

        
    # save
    boxplot_name <- str_replace_all(boxplot_title," ","_")
    plot_and_save(eci_boxplot,
                  title = boxplot_name,
                  height = 5,
                  width = 10)
    
    # check progress
    print(boxplot_name)
  }
}
```

Plot median values of eci
```{r median eci map plots}
for (gu in c("province","country")) {
  bias_tb <- get(paste0(gu,"_individual_bias_eci_stats_HISCO"))
  
  for (bias in c("language","status","gender","agricola","research")) {
    tb <- bias_tb[[bias]]
    if (bias == "language"){
      for (m in paste0("r_",max_distances[1:2])) {
        # extract eci
        eci_m <- tb[[m]]
        # extract stats
        eci_m_median <- eci_m[,c(gu,"median")]
        
        if (gu == "province"){
           map_data <- left_join(
            provinces,
            eci_m_median,
            join_by(province))
        } else if (gu == "country"){
          map_data <- left_join(
            world_sf,
            eci_m_median,
            join_by(iso3==country)) #%>% na.omit()
        }
        map_plot <- ggplot(data = map_data) +
            geom_sf(aes(fill = median)) +
             coord_sf(
                xlim = c(-15,40), 
                ylim = c(20,60)) +
            scale_fill_gradient(low = "white", high = "red", name = "Value") +
            theme_minimal() +
            labs(title = "median eci")
        
        plot_and_save(map_plot,
                      title = paste0(gu,"_",bias,"_bias_correction_",m,"_eci_median_HISCO"))
      } 
    } else {
      # extract median
      eci_median <- tb[,c(gu,"median")]
      
       if (gu == "province"){
           map_data <- left_join(
            provinces,
            eci_median,
            join_by(province))
        } else if (gu == "country"){
          map_data <- left_join(
            world_sf,
            eci_median,
            join_by(iso3==country)) #%>% na.omit() 
        }
        map_plot <- ggplot(data = map_data) +
            geom_sf(aes(fill = median)) +
             coord_sf(
                xlim = c(-15,40), 
                ylim = c(20,60)) +
            scale_fill_gradient(low = "white", high = "red", name = "Value") +
            theme_minimal() +
            labs(title = "median eci")
        
        plot_and_save(map_plot,
                      title = paste0(gu,"_",bias,"_bias_correction_eci_median_HISCO"))
    }
    print(bias)
  }
  print(gu)
}

# original province
 map_data <- left_join(
            provinces,
            province_individual_bias_eci_stats$original$eci,
            join_by(province))
 map_plot <- ggplot(data = map_data) +
            geom_sf(aes(fill = eci)) +
             coord_sf(
                xlim = c(-15,40), 
                ylim = c(20,60)) +
            scale_fill_gradient(low = "white", high = "red", name = "Value") +
            theme_minimal() +
            labs(title = "original eci")
   plot_and_save(map_plot,
                 title = paste0("province","_","original","_eci_HISCO"))
   
# countries
map_data <- left_join(
            world_sf,
            country_individual_bias_eci_stats_HISCO$original$eci,
            join_by(iso3==country)) #%>% na.omit()    
 map_plot <- ggplot(data = map_data) +
            geom_sf(aes(fill = eci)) +
             coord_sf(
                xlim = c(-15,40), 
                ylim = c(20,60)) +
            scale_fill_gradient(low = "white", high = "red", name = "Value") +
            theme_minimal() +
            labs(title = "original eci")
   plot_and_save(map_plot,
                      title = paste0("country","_","original","_eci_HISCO"))
```